---
title: "simpleHarvest"
author: "ieddy & pkalanta"
date: "`r format(Sys.time(), '%d %B %Y')`" #"15 August 2019"
output: pdf_document
            toc: true
            toc_depth: 2
            number_sections: true
            ---
            
            ```{r setup, include=FALSE}
            knitr::opts_chunk$set(
              echo = TRUE,
              message = FALSE,
              warning = FALSE,
              fig.align = "center"
            )
            
library(Require)

Require(c("SpaDES.core", "PredictiveEcology/SpaDES.docs",'RCurl', 'XML', 'snow', 
          'googledrive', 'httr2', "gert", "remotes", "terra", "data.table"),
          upgrade = FALSE, install = FALSE)

# Overview

<!-- Provide an overview of what the module does / how to use the module. -->

<!-- Module documentation should be written so that others can use your module. -->
<!-- This is a template for module documentation, and should be changed to reflect your module. -->

The `simpleHarvest` module simulates spatial forest harvesting within the
LandR framework. It allocates harvest events across the landscape while
attempting to meet annual harvest targets under user-defined constraints.

The module operates on species–age cohort data and management blocks, and
selects eligible pixels based on minimum harvest age, timber harvesting land
base (THLB), and stochastic spread rules. Harvest targets can be provided
externely or calculated dynamically during the simulation.

During each simulation year, `simpleHarvest` generates several outputs for
analysis and visualization, including:

- **`rstCurrentHarvest`**: a raster identifying pixels harvested in the
  current year.  
- **`speciesHarvestMaps`**: species-specific rasters showing harvested
  biomass by species.  
- **`timeSinceHarvest`**: raster tracking the number of years since each
  pixel was last harvested.  
- **`cumulativeHarvestMap`**: raster recording the cumulative harvested area
  or biomass over the simulation period.

`simpleHarvest` is intended as a lightweight and flexible alternative to more
complex harvest models, allowing users to explore management scenarios and
test assumptions about harvest allocation and target-setting.

## R Markdown

R Markdown syntax allows R code, outputs, and figures to be rendered in the documentation.

For help writing in R Markdown, see http://rmarkdown.rstudio.com/.

# Usage

```{r module_usage}
library(SpaDES)
# 1. Set paths
setPaths(modulePath = file.path("/home/ieddy/Git"),
         inputPath = file.path("../../temp"),
         outputPath = file.path("../../temp"),
         cachePath = file.path("../../temp"))

getPaths() # shows where the 4 relevant paths are

# 2. Define simulation times
times = list(start = 2011, end = 2040)

# 3. Define parameters
# parameters <- list(
#   #.progress = list(type = "text", interval = 1), # for a progress bar
#   ## If there are further modules, each can have its own set of parameters:
#   #module1 = list(param1 = value1, param2 = value2),
#   #module2 = list(param1 = value1, param2 = value2)
# )
parameters = list(
    globals = list(
      sppEquivCol = "LandR", 
      .plots = "png",
      .plotInterval = 1,
      .useCache = c(".inputObjects", "init")
    ), 
      simpleHarvest = list(
    .useCache      = ".inputObjects",
    harvestTarget  = c("1" = 0.05, "2" = 0.02, "3"= 0.01, "4"= 0.008)
  )
  # other module parameters can be added similarly
)

# 4. List of modules to load
 modules = c(
    "PredictiveEcology/Biomass_borealDataPrep@development",
    "PredictiveEcology/Biomass_core@development",
    "ianmseddy/LandR_reforestation@parvindev",
    "PredictiveEcology/scfm@development",
    "PredictiveEcology/Biomass_regeneration@development",
    "pkalanta/simpleHarvest@parvintesting"
  ),
 
# 5. Optional pre-existing objects
objects <- list()  
 
inputs <- list()

# 6. Initialize simulation
mySim <- simInit(
  times   = times,
  params  = parameters,
  modules = modules,
  objects = objects
)

 outputs = rbind(
    data.frame(objectName = "cohortData", saveTime = times$start:times$end),
    data.frame(objectName = "pixelGroupMap", saveTime = times$start:times$end)
  ),

mySim <- simInit(times = times, params = parameters, modules = modules,
                 objects = objects)

mySimOut <- spades(mySim)
```

# Events

The `simpleHarvest` module responds to scheduled simulation events and performs
harvest-related actions when triggered. Events are typically scheduled on an
annual timestep.

## Harvest event

The main harvest event is executed once per simulation year. During this event,
the module:

1. Identifies eligible pixels based on minimum harvest age and the timber
   harvesting land base (THLB).
2. Computes block-level harvest targets, either from predefined inputs or
   dynamically based on current landscape conditions.
3. Allocates harvest spatially using a stochastic spread algorithm, subject to
   maximum patch size constraints.
4. Updates cohort ages and biomass for harvested pixels.
5. Records realized harvest statistics for the current year.

This event is responsible for the majority of the module’s functionality and
drives landscape change related to forest harvesting.

## Initialization event (if applicable)

If an initialization event is defined, it is executed at the beginning of the
simulation. This event prepares required data structures, initializes summary
tables, and performs basic validation of inputs.

## Output / summary event (optional)

Some configurations of the `simpleHarvest` module include an output or summary
event, which runs after harvest allocation each year. This event collects
harvest data for later analysis and visualization, including:

- **Annual harvest statistics**, aggregated across blocks and/or species, 
  reporting harvested biomass, area, or pixel counts.  
- **Spatial harvest maps**, such as:
  - `rstCurrentHarvest`: pixels harvested in the current year  
  - `speciesHarvestMaps`: species-specific harvested biomass maps  
  - `timeSinceHarvest`: raster tracking years since last harvest  
  - `cumulativeHarvestMap`: raster of cumulative harvested area or biomass  

These outputs allow users to monitor harvest performance, validate targets,
and visualize both temporal and spatial patterns of harvesting across the
landscape.

## Plotting

The `simpleHarvest` module can produce summary plots to support inspection and
validation of harvest behavior during and after a simulation.

Typical plots include:

- **Annual harvested volume**, showing total realized harvest per year and
  allowing comparison with target harvest levels.
- **Harvested area through time**, summarizing the spatial extent of harvest
  events across simulation years.
- **Block-level harvest summaries**, illustrating how harvest is distributed
  among management blocks.
- **Species-level harvest summaries**, showing the contribution of individual
  species to total harvested volume.

When enabled, spatial plots may also be produced to visualize the location and
configuration of harvested patches for selected years.

These plots are intended for diagnostic and exploratory purposes and can be
used to assess whether harvest targets are met and whether spatial patterns are
consistent with model assumptions.

## Saving

The `simpleHarvest` module saves harvest-related outputs to support downstream
analysis, validation, and visualization.

Saved objects typically include:

- **Annual harvest summary tables**, containing realized harvest volume and/or
  harvested area by year.
- **Block-level harvest statistics**, summarizing harvest allocation within
  management blocks.
- **Species-level harvest statistics**, reporting harvested volume by species
  and year.
- **Spatial harvest outputs** (optional), identifying pixels or patches that
  were harvested in each simulation year.

These outputs are stored within the simulation object for access by subsequent
modules and may also be written to disk depending on user configuration. Saved
outputs are intended to be lightweight and reproducible, enabling users to
reconstruct harvest dynamics without rerunning the full simulation.

# Data dependencies

The `simpleHarvest` module requires the following data inputs to run correctly:

- **`cohortData`** (data.frame / data.table)  
  Contains species–age cohort information for each pixel. Used to identify
  harvestable cohorts, compute biomass, and update post-harvest conditions.  
  **Required.**

- **`pixelGroupMap`** (raster or matrix)  
  Maps each pixel to a management block or group. Used to allocate harvest
  spatially according to block-level rules.  
  **Required.**

- **`thlb`** (raster / logical)  
  Timber harvesting land base, indicating where harvest is allowed. Pixels
  outside the THLB are excluded from harvest.  
  **Required.**

- **`blockId`** (vector or field)  
  Identifiers for management blocks. Used to compute block-level harvest
  targets and summaries.  
  **Required.**

- **`rotationAge`** (list or named vector)  
  Rotation age for each block, used to calculate dynamic harvest targets.  
  **Required.**

- **Optional inputs** may include:
  - **Predefined harvest targets** (if dynamic targets are not used)  
  - **Spatial masks** for inaccessible areas  
  - **Historical harvest maps** for comparison or validation

## Input data

How to obtain input data, and a description of the data required by the module.
If `sourceURL` is specified, `downloadData("simpleHarvest", "path/to/modules/dir")` may be sufficient.

## Output data


The `simpleHarvest` module returns a list containing four main outputs:

### 1. `rstCurrentHarvest`
- **Type:** raster  
- **Contents:** Marks the pixels that were harvested in the current simulation year. Typically, harvested pixels have values > 0, while non-harvested pixels are `NA`.  
- **Purpose:** Useful for spatial visualization of harvest events and integration with other landscape analyses.

### 2. `speciesHarvestMaps`
- **Type:** list of rasters (one per species)  
- **Contents:** For each species, maps of harvested pixels showing species-specific biomass removal.  
- **Purpose:** Allows analysis of species-level harvest impacts across the landscape.

### 3. `harvestStats`
- **Type:** data.table / data.frame  
- **Contents:** Tabular summary of harvest events for the current year, including:
  - `blockId` or pixel group
  - `speciesCode`
  - `harvested biomass` or area  
- **Purpose:** Provides block- and species-level summaries for reporting and validation.

### 4. `harvestPerformance`
- **Type:** data.table / data.frame  
- **Contents:** Tracks how well the module met harvest targets, including:
  - planned vs realized harvest
  - discrepancies relative to feasible maximum  
- **Purpose:** Used for validation, performance evaluation, and debugging.

# Links to other modules

The `simpleHarvest` module interacts with other modules through its input
requirements and outputs:

### Input linkages
- **Forest succession / cohort dynamics modules**: provide `cohortData`
  required to identify harvestable cohorts.  
- **Landscape / landbase modules**: provide `thlb` and `pixelGroupMap` to
  define harvestable areas and management blocks.  

### Output linkages
- **Habitat modules**: use `rstCurrentHarvest` and `speciesHarvestMaps` to
  update habitat quality for wildlife.  
- **Carbon or biomass modules**: use `harvestStats` and
  `harvestPerformance` to quantify biomass removal or carbon flux.  
- **Visualization modules**: use spatial outputs to generate maps of harvested
  areas.  

These linkages allow `simpleHarvest` to integrate with the simulation workflow,
ensuring that harvest events influence downstream ecological processes and
that the module can be informed by upstream landscape dynamics.

